Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-10 edubots_surveys.Rnw
Review again methods, all pars

This study has employed a quantitative approach in order to address its research objectives. A good amount of data was collected on the usage of messaging applications and chatbots by educators in universities and colleges.\footnote{In Spain, higher education institutions generally known elsewhere as colleges, are called simply high schools, and deliver
tertiary or associate degrees that do not have the consideration of university degrees.}
To achieve this goal two online surveys were designed and developed using Google Forms
in Spanish. They are comprised of six questions focusing on demographic data
(i.e., sector, gender, degrees, discipline, age, or teaching experience)
followed by several multiple choice questions allowing the participants to choose multiple answers.

In the {\bf first survey}, multiple choice questions were
focused on the use of messaging apps in the teaching practice, the type of chatbot use cases that educators would consider
useful for their teaching and the impact of COVID on the teaching practice; please, check them out to the full extent in the Appendix.\footnote{This {\bf first survey} was made in two separate Google forms, one for
university and another for teachers in tertiary education institutions, outside the university system (vocational training, including 1- and 2-year degrees). Questions and
responses were the same, except for the type of tertiary degrees that
were considered.} This survey was piloted by the authors of this paper and their colleagues before using it in the
study. All the comments collected from this trial were incorporated in the survey.
We primarily used mailing lists and Telegram groups to disseminate the survey link so that it would reach a larger number of educators. The form targeting university teachers was sent to Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also universities in Costa Rica and
Mexico. The tertiary (non-university/college) educators who received the form were based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, during the Covid-19 pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.

A total of 282 educators responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). From those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%) and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of participants (n=111) in this survey were 45-55 years old (39.4\%), while 91 educators were 35-45 (32.3\%), 46 were 25-35 (16.3\%) and 34 were older than 55 (12.1\%). Last, 84 educators had a teaching experience of 16-25 years (29.8\%), 75 educators had 6-15 years of experience (26.6\%), 69 educators had 0-5 years of experience (24.5\%) and finally, 54 educators had more than 25 years of teaching experience (19.1\%).

Responses were stored automatically in a Google Drive spreadsheet. Eventually,
the results from the two forms we used for the first survey were collated in a
single spreadsheet. Survey questions are shown in Appendix I (Section \ref{app:questions}).

The {\bf second survey} was designed after initial results for the first survey arrived,
and pointed to necessities and experiences of educators not addressed by the first one; namely
their experience with messaging platforms and the way they were used to interact with students.
It was piloted in a group of university teaching staff attending to a formation course, and validated with them.
It was then extended to the rest of responders, using the same media: Telegram groups, e-mail, and announcement in mailing lists. The questions asked in that survey are also shown in Appendix I
(Section \ref{app:questions}).

A total of 205 educators responded to the second survey: 187 graduate teachers (91.2\%) and 18 student teachers (8.8\%). From those, 124 were male (60.5\%), 65 female (31.7\%) and 16 participants preferred not to indicate their gender (7.8\%). In terms of age, the majority of the participants (n=70) was again, as in the first survey, 45-55 years old (34.1\%), while 67 participants were 35-45 years old (32.7\%), 42 were 25-35 (20.5\%) and 26 participants were older than 55 (12.7\%). In terms of teaching experience, 59 participants had a teaching experience of 16-25 years (28.8\%), 51 participants had 6-15 years (24.9\%) and 0-5 years of experience (24.9\%), and 44 participants had more than 25 years of teaching experience (21.5\%).

>>>>

This study employed a quantitative approach to address its research objectives. A significant amount of data was collected on the usage of messaging applications and chatbots by educators in universities and colleges.\footnote{In Spain, higher education institutions that would generally known as colleges in other places, are simply called high schools, and provide
tertiary or associate degrees that do not have the same consideration as university degrees.}
To collect this data, two online surveys were designed and developed using Google Forms
in Spanish. These surveys consisted of six questions on demographic data
(such as sector, gender, degrees, discipline, age, or teaching experience)
followed by several multiple-choice questions allowing participants to choose multiple answers.

The {\bf first survey} consisted of multiple-choice questions 
focused on the use of messaging apps in teaching practice, the type of chatbot use cases that educators would find
useful for their teaching and the impact of COVID on teaching practice. These questions can be found in their entirity in the Appendix.\footnote{This {\bf first survey} was split into two separate Google forms, one for
university teachers and another for teachers in tertiary education institutions outside the university system (vocational training, including 1- and 2-year degrees). The questions and
responses were the same, except for the type of tertiary degrees that
were considered.} This survey was piloted by the authors of this paper and their colleagues before using it in the
study, and all feedback from this trial was incorporated into the survey.
The survey link was primarily distributed via mailing lists and Telegram groups in order to reach a larger number of educators. The form targeting university teachers was sent to university in Spain
(mainly in Andaluc\'{\i}a and Galicia), and also universities in Costa Rica and
Mexico. Tertiary (non-university/college) educators who received the form were mainly based in
Andaluc\'{\i}a. The survey was disseminated and responses were collected in the first quarter of
2021, during the Covid-19 pandemic, while at least in Spain, many universities were implementing
mandatory virtual teaching.

A total of 282 educators responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). Of those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%), and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of participants (n=111) in this survey were 45-55 years old (39.4\%), while 91 educators were 35-45 (32.3\%), 46 were 25-35 (16.3\%), and 34 were older than 55 (12.1\%). Lastly, 84 educators had a teaching experience of 16-25 years (29.8\%), 75 educators had 6-15 years of experience (26.6\%), 69 educators had 0-5 years of experience (24.5\%), and finally, 54 educators had more than 25 years of teaching experience (19.1\%).

Responses were automatically stored in a Google Drive spreadsheet. Eventually,
the results from the two forms used for the first survey were collated in a
single spreadsheet. Survey questions can be found in Appendix I (Section \ref{app:questions}).

The {\bf second survey} was designed after initial results for the first survey arrived,
and pointed out necessities and experiences of educators not covered by the first survey, specifically
their experience with messaging platforms and the ways they were used to interact with students.
It was piloted with a group of university teaching staff attending a training course, and validated by them.
The survey was then extended to the rest of responders, using the same media: Telegram groups, email, and announcements in mailing lists. The questions asked in this survey can also be found in Appendix I
(Section \ref{app:questions}).

A total of 205 educators responded to the second survey: 187 graduate teachers (91.2\%) and 18 student teachers (8.8\%). Of those, 124 were male (60.5\%), 65 female (31.7\%), and 16 participants preferred not to indicate their gender (7.8\%). In terms of age, the majority of the participants (n=70) were again, as in the first survey, 45-55 years old (34.1\%), while 67 participants were 35-45 years old (32.7\%), 42 were 25-35 (20.5\%), and 26 participants were older than 55 (12.7\%). In terms of teaching experience, 59 participants had a teaching experience of 16-25 years (28.8\%), 51 participants had 6-15 years (24.9\%), 51 had 0-5 years of experience (24.9\%), and 44 participants had more than 25 years of teaching experience (21.5\%).
